Self-sustained carbon monoxide oxidation oscillations on size-selected platinum nanoparticles at atmospheric pressure

### 2012-04-16 paper.tex
Started integrating Ibs introduction into main document

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
CO oxidation is one of the most studied reactions due to its importance in both automotive catalysis and in the removal of trace CO in hydrogen streams. A large effort and a large number of experiments with a variety of surface sensitive techniques have been conducted on single crystal surfaces greatly improving the understanding of this reaction. Single crystal surfaces are well suited model systems due to the high degree of achievable control of parameters compared to industrial catalysts. Here we report on studies in a microreactor model system of Pt nanoparticles. The size-selected nanoparticles supported on a SiO$_2$ substrate show oscillatory behavior of CO oxidation at atmospheric pressure.

Oscillating reactions on single crystals at UHV conditions have been studied extensively and is know to originate from surface reconstructions of the surface during the oscillations\cite{Ertl2008}. It has previously been shown\cite{SALES1982} that CO oxidation at atmospheric pressures also shows oscillatory behavior on single crystals of Pt, Pd and Ir. Recent studies have proposed a mechanism on Pd \cite{Hendriksen2010} attributed to a cycle of surface oxide formation, oxide roughening and clean metal surface formation. CO oxidation oscillations has also been observed on Pt thin films\cite{Lund2000}, in macroscopic flow reactors\cite{Singh2010} and on Pt nanoparticles synthesized from wet chemical synthesis and subsequent annealing \cite{Carlsson2006}. Here we investigate CO oscillations in a microreactor model system on size selected Pt nanoparticles by high sensitivity mass spectrometry.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Self-sustained oscillating systems are wide spread in nature and have been
described both experimentally and theoretically. One of the maybe most famous
biological examples is the oscillation in the lynx-hare population as
reported through the number furs delivered to the Hudson Bay Company A).
This was for some time was believed to be forced by the sunspot oscillations,
however, this proved to be a coincidence B) and could be described alone
by a close dynamical coupling in between the two populations. Similar very
interesting observations have been made for our current climate changes
\cite{Svensmark1997,Svensmark1998} suggesting that the sunspot activity could
influences the level of cosmic rays and thus the cloud formation, but this
currently a topic of controversy \cite{Lockwood2007}. A beautiful example of
self-sustained oscillatory behavior in chemistry is the colorful
Belousov-Zhabotinsky reaction which is rather easy to demonstrate for an
audience\cite{ZAIKIN1970}. Such self-sustained oscillatory behavior has been
explained for systems far from equilibrium in several theoretical works
\cite{HakenBog,NicolisBog}. Similar behavior has been observed within
heterogeneous catalysis where typically nanoparticles of a precious metal
supported on an inexpensive oxide constitute the active phase \cite{IbsBog}.
One of the very first examples was observed by Wicke for CO oxidation on a
catalyst of highly dispersed Pt on γ−Al2O3 \cite{BEUSCH1972}. To gain a deeper
insight into the fundamental aspects of such oscillatory reactions this work
was followed up by Ertl on Pt(110) surfaces under well-defined conditions
\cite{EISWIRTH1986}. Ertl demonstrated how the switching in between the
unreconstructed Pt(110) surface and the reconstruction 2x1-Pt(110) surface
combined with big changes of oxygen sticking coefficient on the two surfaces
could drive the self-sustained oscillations. This and many other interesting
aspects of self-sustained and forced oscillatory behavior and spatio–temporal
self-organization were investigated by Ertl and coworkers on well-defined
single crystals see for example reviews \cite{IMBIHL1995,JAKUBITH1990} and in
particular the Nobel Lecture \cite{Ertl2008}. One of the reasons this has
attracted so much interest is naturally the importance of such catalytic
reactions and therefore a demand to gain a better understanding of their
behavior with the goal of further improvement. CO oxidation is one of the most
studied reactions due to its importance in both automotive catalysis and in
the removal of trace CO in hydrogen streams \cite{IbsBog}. The oscillatory
behavior is, however, limited to either Pt or the CO oxidation, but
have been found for many other metallic catalyst like on for example Rh, Ir,
and Pd and for oxidations reactions in general as well as for NO and CO
reactions and even for hydrogenation reactions \cite{Lund2000,SALES1982}. Most
of the studies are made on single crystal or poly crystalline films or
filaments, mainly in conjunction with ultra-high vacuum equipment so that the
surfaces can be characterized using the surface science methods. The amount of
work described on supported catalyst seems more limited obviously because it
is so much more difficult to characterize those systems and because additional
complexities are present such as mass and heat transport. There even exist
reports in the literature of how large scale facilities such as an industrial
ammonia fixed-bed synthesis reactor in Germany in 1989 started oscillating for
several hours. This could be modeled by a coupling of the exothermic process
and the heat management in the reactor system \cite{Machado2005}, but one
cannot rule out that the observed oscillations could in some systems also be
created by inherent processes on the catalyst itself. In general such
oscillations would be considered highly undesirable in industry since they
could be difficult to control and result in run-away incidents in particular
for exothermic reactions resulting in damaged catalyst due to sintering at
higher temperatures. On the other hand there exist examples in the literature
where forced oscillations seem to improve the overall activity
\cite{IMBIHL1995,Machado2005} . One particular interesting process here could
be the methanol synthesis where reducing the CO$_2$ contents for a shorter period
of time result in a subsequent higher methanol rate over compensation for the
lack of methanol production when running in the reducing H$_2$ and CO alone
\cite{Dynamics-Bog}. This enhancement can be explained by extensive wetting
and un-wetting of the active Copper catalyst dependent on the oxidation
potential \cite{Hansen2002} i.e. the active area is strongly increased when
reduced but slowly returns to the less active state when the CO$_2$ is present.

As mentioned the understanding of self-sustained on supported catalyst is
hampered by the fact that entirely different mechanism than those observed on
single and polycrystalline samples may exist. For example on a Pt(110) crystal
it was shown how waves of surface reconstruction could propagate on the
surface ensuring a synchronization of the phenomena. This also have to occur
for oscillation on supported catalyst, but here it is more likely to be due to
coupling of mass or heat transport. There have been a few examples where such
oscillations have been observed. Of these are the most prominent the work by
Skoghlund \cite{Carlsson2006} where self-sustained oscillations were observed
on supported Pt catalyst and simultaneous studied by in situ IR spectrometry.
Similar studies were recently undertaken using EXAFS \cite{Singh2010} but here
the oscillatory behavior was rather unstable and somewhat less convincible.

In this work we demonstrate how a two-dimensional model system of Pt
nanoparticles on a SiO2 support can be constructed that displays highly
repeatable self-sustained oscillations when first started. The system consist
of mass selected Pt nanoparticles in the range from 2 to 10\,nm Pt
nanoparticles deposited in a microreactor with a surface area of 0.75\,cm$^2$
specially designed to achieve high sensitivity \cite{Henriksen2009}. This
method offers some advantages over the chemical synthesis methods, which also
offers the possibility of size selected nanoparticles se for example 
\cite{Tsang2008,Shao2011}, since in this case we do not have to bother with
the surfactants required for these synthesis methods. The coverage of Pt
nanoparticles could be varied from sub-percent to full coverage. It is
demonstrated how the nanoparticles could provide self-sustained oscillations
over extend of days and how various parameters could be mapped out inclusive
the influence of the temperature dependence - an effect that could not be
demonstrated for a continuous film of same area and similar conditions. By
combining these results with previous studies \cite{Carlsson2006,Singh2010}
and a recent detailed study of CO oxidation on Pd using STM
\cite{Hendriksen2010} a plausible explanation of the oscillatory behaviors
presented. The low coverage and extremely efficient Pt nanoparticles
illuminates some of the pit falls that may have to be considered when making
in situ investigations of such a system since extremely low coverages may
provide the entire conversion.

A) D. A. McLulich, Variations in the Number of Varying Hare,University of Toronto
Press, Toronto, 1937. (we need to dig in for the correct ref here),
B) I remember reading of this in a biology book back in the seventies, check)
